Using Machine Learning and Constraint Programming to Solve the Relational-To-Ontology Mapping Problem

### 2017-08-11 stp/paper_aaai/main.tex
alignment graph

\usepackage{tikz}
\usetikzlibrary{shadows.blur}
\usetikzlibrary{shapes.symbols}
\usetikzlibrary{arrows,shapes,decorations.pathmorphing}
\usetikzlibrary{positioning}
\usetikzlibrary{decorations.text}
\usetikzlibrary{decorations.pathmorphing}
\tikzstyle{class}=[draw,ellipse,align=center,fill=green!25]
\tikzstyle{data}=[draw,rectangle,align=center,black!70]
\tikzstyle{hasaA} = 
[draw,densely dashed,-,black!70,decoration={snake,segment 
length=10,amplitude=1.5,post length=4}]
\tikzstyle{hasaC} = [>=stealth',draw,thick,-, black]
\tikzstyle{isa} = [>=latex,draw,double,black!90,double distance=1mm]

@@

\end{tikzpicture}
\caption{Example of ontology
(`\tikz{ \node[class] () at (0,0)  {};}' are Classes, `\tikz{ \path[hasaC,->] 
(0,1) -- (3ex,1);}' means ``has a Class property'', 
`\tikz{ \path[hasaA,decoration={post length=0}] (0,0) edge[decorate] (3ex,0);}' 
means ``has 
a data 
property'' and 
`\tikz{ \path[isa,->] (0,1) -- (3ex,1);}' means ``is a'' ).
}
\label{FIG:onto}
\end{figure}

@@

\begin{figure}[ht]
\centering
\begin{tikzpicture}
\begin{scope}
	\node[class] (pe) at (0,0)  {\small Person};
	\node[class] (co) at (6,0) {\small Org.};
	\node[class] (ci) at (2.5,0) {\small City};
	\node[class] (st) at (4.5,0) {\small State};

	\node[data] (pen) at (-1,-1) {\scriptsize Person.name};
	\node[data] (peb) at (1,-1) {\scriptsize Person.birthDate};
	\node[data] (cin) at (3,-1) {\scriptsize City.name};
	\node[data] (stn) at (4.5,-1) {\scriptsize State.name};
	\node[data] (con) at (6,-1) {\scriptsize Org.name};

	\path[hasaC,->] (pe) edge [bend left= 20] 
	node [midway, above, sloped] {worksFor} (co) ;

@@

\end{scope}
\end{tikzpicture}
\caption{Examples of two semantic models.}
\label{FIG:sem}
\end{figure}

@@

We formulate the problem of semantic labeling as a multi-class classification
problem. The known source descriptions $S_T$ provide us the training sample. We
compute a feature vector for each source \ddg{Which features?} attribute and 
associate the known
semantic type with the corresponding feature vector. We train random forest
on this sample. Then we use the learnt model to classify attributes in the new
source.

In such way, we learn the mapping $\psi : \mathcal{A_s} \times \mathcal{L_O} 
\mapsto [0, 1]$,
where $\psi(a_i,l_j)$ indicates the confidence of the attribute $a_i$ to be 
mapped to the semantic type $l_j$.

@@

An example alignment graph is illustrated in Fig. \ddg{todo}. Black links 
correspond to the links which are supported by the known semantic models. Blue
links are inferred from the ontology $\mathcal{O}$ and can be assigned the 
maximum weight $w_max$ \ddg{where does this weight come from?} .

@@

Let $G = (V_G,E_G,\lambda_G)$ and $P = (V_P,E_P, \lambda_P)$ be two graphs 
where $\lambda_P$ is a \emph{labeling function} that maps nodes/edges of $G$ 
into a set of labels (e.g. natural numbers) $\mathbb{N}$ (similarly for 
$\lambda_P$).
An \emph{embedding} of $P$ into $G$ is an injective function $f : V_G \mapsto 
V_P$ such that for all $x,y \in V_P$:
\begin{enumerate}
	\item $\{x,y\} \in E_P \implies \{f(x),f{y}\} \in E_G$.
	\item $\lambda_P(x) = \lambda_G(f(x))$
	\item $\lambda_P(\{x,y\}) = \lambda_G(\{f(x),f(y)\})$
\end{enumerate} 
That is, if an edge exists in $P$, it must also exist in $G$ (but not 
vice-versa); and the labelling functions of both graphs must match after 
applying the embedding function.

@@

Given a graph $G =
(V, E)$ and a subset of its nodes $T \subseteq V$ , called \emph{terminals}, a 
Steiner Tree $G_s = (V_s, E_s )$ is a tree such that $T \subseteq V_s \subseteq 
V$ and $E_s \subseteq E$. In other words, $G_s$ spans all the nodes in $T$ and 
may include additional nodes from $V$, in particular to ensure the 
connectedness of the constructed tree. The Steiner Tree Problem 
(STP) is stated as follows: given $G$ and a weight function $w_f : E 
\longmapsto \mathbb{R}$, find the Steiner Tree that minimizes the sum of the 
weights of the edges in $E_s$ given by $w_f$. This was proven to be NP-hard 
by Karp \cite{Karp1972}.

To formulate the \relonto{} schema mapping problem as a STP for a new source 
$s^\star$, we construct the 
\emph{integration graph} $\mathcal{I}_\mathcal{O}^{s^\star} = 
(\mathcal{V}_\mathcal{O}^{s^\star},\mathcal{E}_\mathcal{O}^{s^\star})$, with 
nodes $\mathcal{V}_\mathcal{O}^{s^\star} = \mathcal{V_O} \cup 
\mathcal{A}_{s^\star}$, 
edges $\mathcal{E}_\mathcal{O}^{s^\star}$.

The set of edges $\mathcal{E}_\mathcal{O}^{s^\star}$ is constructed by using 
all the edges in the alignment 
graph, and edges connecting each attributes of $s^\star$ to the nodes 
$\mathcal{D_{G_O}}$ of the 
alignment graph \emph{induced} by the semantic labels. We call this last set 
of edges $\mathcal{M}_\mathcal{O}^{s^\star}$ (for ``matches''). Thus, 
$\mathcal{E}_\mathcal{O}^{s^\star} = 
\mathcal{E_O} \cup \mathcal{M}_\mathcal{O}^{s^\star}$.

@@

Note that, although the alignment graph is directed, the semantic models are 
not necessarily rooted directed trees (e.g. \ref{FIG:sem}), therefore the 
directions of the edges is ignored for the STP model (and are all treated as 
undirected edges).

@@

The solution $T^\star$ must satisfy the following constraints:
\begin{enumerate}
	\item $T^\star$ must be a subgraph of $\mathcal{I}_\mathcal{O}^{s^\star}$.
	\item $T^\star$ must be a tree (connected acyclic graph).
	\item $\forall a \in \mathcal{A}_{s^\star}, a\in V^\star$.
	\item $\forall a \in \mathcal{A}_{s^\star}, degree(a) = 1$.
	\item $\forall n \in \mathcal{D_{G_O}}, degree(a) \in 
	\{0,2\}$
	\item $\forall n \in \mathcal{D_{G_O}} \cap V^\star, 
	degree(a) = 2$
\end{enumerate}

@@

\subsection{Modeling \relonto{} in CP}
In order to model this problem we used the \minizinc{} language, and the 
\chuffed{} solver because it has a global constraint implemented for Steiner 
Tree Problems \cite{deuna2016steiner}.

Because attributes must be connected to exactly one node of the alignment 
graph, and that node will be in $\mathcal{D_{G_O}} \cap V^\star$ 
then the part of the problem between attribute nodes and the alignment graph is 
actually a matching problem. Each attribute must match exactly to one node of 
$\mathcal{D_{G_O}}$.

@@

\begin{flalign}
	& \text{Minimize~~} w_{STP} + w_{ADIFF} - w_{PAT}
	%\sum_{e \in \mathcal{E}_\mathcal{O}}c_e * 
	%w_\mathcal{I}(e) + \sum_{e \in \mathcal{M}_\mathcal{O}^{s^\star}} c_e * 
    %		w_\mathcal{I}(e)
	\label{EQ:obj}&&\\
	&\text{such that~~}  \nonumber&& \\
	& \textit{steiner\_tree}(\{c_n| n \in \mathcal{V_O}\},\{c_e| e \in 
	\mathcal{E_O}\},\mathcal{G_O},w_\mathcal{O},w_{STP})  \label{EQ:stp} &&\\
	&\forall d \in \mathcal{D_{G_O}}, degree(d) \leq 1 \label{EQ:deg1}&&\\
	&\forall d \in \mathcal{D_{G_O}}, c_d \Leftrightarrow degree(d) = 1 
	\label{EQ:deg2}&&\\
	&\forall a \in \mathcal{A}_{s^\star}, match[a] \in \{ d | (a,d)\in 
	\mathcal{M}_\mathcal{O}^{s^\star} \} \label{EQ:matchdom}&&\\
	& \textit{alldifferent}(match) \label{EQ:alld}&& \\
	& \forall a \in \mathcal{A}_{s^\star}, c_{match[a]} = \mathit{true} 
	\label{EQ:map}&&\\
	& w_{\mathit{ADIFF}} = \sum_{(a,d) \in \mathcal{M}_\mathcal{O}^{s^\star}} 
	w_\mathcal{I}(~(a,d)~) * \llbracket match[a] = d\rrbracket 
	\label{EQ:matchcost}  &&\\
	& \forall p \in \mathcal{P}, \big(\forall e\in edges(p), c_e = 
	\mathit{true} \big) \Leftrightarrow c_p \label{EQ:patt}&&\\
	& w_{PAT} = \sum_{p\in\mathcal{P}} \mathit{support}[p] * c_p 
	\label{EQ:pcost} 
\end{flalign}

>>>>

\usepackage{tikz}
\usetikzlibrary{shadows.blur}
\usetikzlibrary{shapes.symbols}
\usetikzlibrary{arrows,shapes,decorations.pathmorphing}
\usetikzlibrary{positioning}
\usetikzlibrary{decorations.text}
\usetikzlibrary{decorations.pathmorphing}
\usetikzlibrary{mindmap,backgrounds}
\tikzstyle{class}=[ellipse,align=center,fill=green!25]
\tikzstyle{data}=[draw,rectangle,align=center,black!70]
\tikzstyle{hasaA} = 
[draw,densely dashed,-,black!70,decoration={snake,segment 
length=10,amplitude=1.5,post length=4}]
\tikzstyle{hasaC} = [>=stealth',draw,thick,-, black]
\tikzstyle{isa} = [>=o,draw,densely dotted,very thick,black!90]

@@

\end{tikzpicture}
\caption{Example of ontology
(`\tikz{ \node[class] () at (0,0)  {};}' are Classes, `\tikz{ \path[hasaC,->] 
(0,1) -- (3ex,1);}' means ``has a Class property'', 
`\tikz{ \path[hasaA,decoration={post length=0}] (0,0) edge[decorate] (3ex,0);}' 
means ``has 
a data 
property'' and 
`\tikz{ \path[isa,->] (0,1) -- (3ex,1);}' means ``is a'' ). Based on 
\cite{Taheriyan2013}.
}
\label{FIG:onto}
\end{figure}

@@

\begin{figure}[ht]
\centering
\begin{tikzpicture}
\begin{scope}
	\node[class] (pe) at (3,0)  {\small Person};
	\node[class] (co) at (6,0) {\small Org.};
	\node[class] (ci) at (1,0) {\small City};
	\node[class] (st) at (-1,0) {\small State};

	\node[data] (pen) at (2.5,-1) {\scriptsize Person.name};
	\node[data] (peb) at (4.3,-1) {\scriptsize Person.birthDate};
	\node[data] (cin) at (1,-1) {\scriptsize City.name};
	\node[data] (stn) at (-1,-1) {\scriptsize State.name};
	\node[data] (con) at (6,-1) {\scriptsize Org.name};

	\path[hasaC,->] (pe) edge [bend left= 0] 
	node [midway, above, sloped] {worksFor} (co) ;

@@

\end{scope}
\end{tikzpicture}
\caption{Examples of two semantic models (`\tikz{ \node[class] () at (0,0)  
{};}' are class nodes, and `\tikz{ \node[data] () at (0,0)  {};}' are data 
nodes).}
\label{FIG:sem}
\end{figure}

@@

We formulate the problem of semantic labeling as a multi-class classification
problem. The known source descriptions $S_T$ provide us the training sample. We
compute a feature vector for each source \ddg{Which features?} attribute and 
associate the known
semantic type with the corresponding feature vector. We train a random forest
on this sample. Then we use the learnt model to classify attributes in the new
source.

In such way, we learn the mapping $\psi : \mathcal{A}_s \times \mathcal{L_O} 
\mapsto [0, 1]$,
where $\psi(a_i,l_j)$ indicates the confidence of the attribute $a_i$ to be 
mapped to the semantic type $l_j$.

@@

An example alignment graph is illustrated in Fig. \ref{FIG:ali}. Bigger 
examples 
can be found in \cite{Taheriyan2013}. Black links 
correspond to the links which are supported by the known semantic models. Blue 
and orange links are inferred from the ontology $\mathcal{O}$ (the orange ones 
correspond to the ones added to connect Classes od the ontology that never 
appeared in the semantic models) and can be assigned the 
maximum weight $w_max$ \ddg{where does this weight come from?} .

\begin{figure}[ht]
\centering
\begin{tikzpicture}

	\node[class] (pe1) at (3,0)  {\small Person};
	\node[class] (co1) at (6,0) {\small Org.};
	\node[class] (ci1) at (1,0) {\small City};
	\node[class] (st1) at (-1,0) {\small State};

	\node[data] (pen) at (2.5,1) {\scriptsize Person.name};
	\node[data] (peb) at (4.3,1) {\scriptsize Person.birthDate};
	\node[data] (cin) at (1,1) {\scriptsize City.name};
	\node[data] (stn) at (-1,1) {\scriptsize State.name};
	\node[data] (con) at (6,1) {\scriptsize Org.name};

	\path[hasaC,->] (pe1) edge [bend left= 0] 
	node [midway, above, sloped] {worksFor} (co1) ;

	\path[hasaC,->] (pe1) edge [bend left = 0] 
	node [midway, above, sloped] {bornIn} (ci1) ;

	\path[hasaC,->] (ci1) edge  []
	node [midway, above, sloped] {state} (st1) ;

	\path[draw,hasaA,->] (pe1) -- (pen);
	\path[draw,hasaA,->] (pe1) -- (peb);

	\path[draw,hasaA,->] (co1) -- (con);

	\path[draw,hasaA,->] (ci1) -- (cin);

	\path[draw,hasaA,->] (st1) -- (stn);

% % % CC2
\def\ysh{-2.5};

	\node[class] (pl2) at (-1,0+\ysh)  {\small City};
	\node[class] (pe2) at (3,0+\ysh)  {\small Person};
	\node[class] (co2) at (1,0+\ysh) {\small Org.};
	\node[class] (ev2) at (6,0+\ysh) {\small Event};

	\node[data] (pen) at (2.5,-1+\ysh) {\scriptsize Person.name};
	\node[data] (evs) at (4.2,-1+\ysh) {\scriptsize Event.startDate};
	\node[data] (eve) at (6,-1+\ysh) {\scriptsize Event endDate};
	\node[data] (con) at (1,-1+\ysh) {\scriptsize Org.name};
	\node[data] (pln) at (-1,-1+\ysh) {\scriptsize City.name};

	\path[hasaC,->] (co2) edge [bend left= 0] 
	node [midway, above, sloped] {ceo} (pe2) ;

	\path[hasaC,->] (ev2) edge [bend left = 0] 
	node [midway, above, sloped] {organizer} (pe2) ;

	\path[hasaC,->] (co2) edge [bend left = -10] 
	node [midway, above, sloped] {location} (pl2) ;

	\path[draw,hasaA,->] (pl2) -- (pln);

	\path[draw,hasaA,->] (pe2) -- (pen);

	\path[draw,hasaA,->] (co2) -- (con);

	\path[draw,hasaA,->] (ev2) -- (evs);
	\path[draw,hasaA,->] (ev2) -- (eve);

% % % Ontology

	\path[draw,hasaC,->,blue!70] (co2)  edge [bend left = -5] (pe1);
	\path[draw,hasaC,->,blue!70] (pe1)  edge [bend left = -5] (co2);

	\path[draw,hasaC,->,blue!70] (co1)  edge [bend left = -5] (pe2);
	\path[draw,hasaC,->,blue!70] (pe2)  edge [bend left = -5] (co1);

	\path[draw,hasaC,->,blue!70] (pl2) -- (st1);

	\path[draw,hasaC,->,blue!70] (ev2) -- (pe1);

	\node[class,fill=orange!30!white] (PL) at (3,\ysh/2)  {\small Place};

	\path[draw,hasaC,->,orange!30!white] (pe1) edge[bend left=10] (PL);
	\path[draw,hasaC,->,orange!30!white] (pe1) edge[bend left=-10] (PL);

	\path[draw,hasaC,->,orange!30!white] (pe2) edge[bend left=10] (PL);
	\path[draw,hasaC,->,orange!30!white] (pe2) edge[bend left=-10] (PL);

	\path[draw,hasaC,->,orange!30!white] (co1) edge[bend left=0] (PL);
	\path[draw,hasaC,->,orange!30!white] (co2) edge[out=45,in=225] (PL);

	\path[draw,hasaC,->,orange!30!white] (ev2) edge[out=90,in=0] (PL);

	\path[draw,isa,->,orange!30!white] (ci1) edge[out=-45,in=120] (PL);
	\path[draw,isa,->,orange!30!white] (pl2) edge[out=90,in=180] (PL);
	\path[draw,isa,->,orange!30!white] (st1) edge[out=-45,in=150] (PL);

\end{tikzpicture}
\caption{Example of alignment graph.}
\label{FIG:ali}
\end{figure}

@@

Let $G = (V_G,E_G,\lambda_G)$ and $P = (V_P,E_P, \lambda_P)$ be two graphs 
where $\lambda_G$ (resp. $\lambda_P$) is a \emph{labeling function} that maps 
nodes/edges of $G$ (resp. $P$) into a set of labels (e.g. natural numbers).
An \emph{embedding} of $P$ into $G$ is an injective function $f : V_G \mapsto 
V_P$ such that for all $x,y \in V_P$:
\begin{enumerate}
	\item $\{x,y\} \in E_P \implies \{f(x),f{y}\} \in E_G$.
	\item $\lambda_P(x) = \lambda_G(f(x))$
	\item $\lambda_P(\{x,y\}) = \lambda_G(\{f(x),f(y)\})$
\end{enumerate} 
That is, if an edge exists in $P$, it must also exist in $G$ (but not 
vice-versa); and the labelling functions of both graphs must match after 
applying the embedding function.

@@

Given a graph $G =
(V, E)$ and a subset of its nodes $T \subseteq V$ , called \emph{terminals}, a 
Steiner Tree $G_s = (V_s, E_s )$ is a tree such that $T \subseteq V_s \subseteq 
V$ and $E_s \subseteq E$. In other words, $G_s$ spans all the nodes in $T$ and 
may include additional nodes from $V$, in particular to ensure the 
connectedness of the constructed tree. The Steiner Tree Problem 
(STP) is stated as follows: given $G$ and a weight function $w_f : E 
\mapsto \mathbb{R}$, find the Steiner Tree that minimizes the sum of the 
weights of the edges in $E_s$ given by $w_f$. This was proven to be NP-hard 
by Karp \cite{Karp1972}.

To formulate the \relonto{} schema mapping problem as a STP for a new source 
$s^\star$, we construct the 
\emph{integration graph} $~\mathcal{I}_\mathcal{O}^{s^\star} = 
(\mathcal{V}_\mathcal{O}^{s^\star},\mathcal{E}_\mathcal{O}^{s^\star})$, with 
nodes $\mathcal{V}_\mathcal{O}^{s^\star} = \mathcal{V_O} \cup 
\mathcal{A}_{s^\star}$, 
edges $\mathcal{E}_\mathcal{O}^{s^\star}$.

The set of edges $\mathcal{E}_\mathcal{O}^{s^\star}$ is constructed by using 
all the edges in the alignment 
graph, and edges connecting each attributes of $s^\star$ to the nodes in the 
alignment graph \emph{induced} by the semantic types (i.e. the set of nodes in 
$\mathcal{D_{G_O}}$). We call this last set 
of edges $\mathcal{M}_\mathcal{O}^{s^\star}$ (for ``matches''). Thus, 
$\mathcal{E}_\mathcal{O}^{s^\star} = 
\mathcal{E_O} \cup \mathcal{M}_\mathcal{O}^{s^\star}$.

@@

Note that, although the alignment graph is directed, the semantic models are 
not necessarily rooted directed trees (e.g. Fig. \ref{FIG:sem}), therefore the 
direction of the edges is ignored for the STP model (and are all treated as 
undirected edges).

@@

The solution $T^\star$ must satisfy the following constraints:
\begin{enumerate}
	\item $T^\star$ must be a subgraph of $\mathcal{I}_\mathcal{O}^{s^\star}$
	\item $T^\star$ must be a tree (connected acyclic graph)
	\item $\forall a \in \mathcal{A}_{s^\star}, a\in V^\star$
	\item $\forall a \in \mathcal{A}_{s^\star}, degree(a) = 1$
	\item $\forall n \in \mathcal{D_{G_O}}, degree(a) \in 
	\{0,2\}$
	\item $\forall n \in \mathcal{D_{G_O}} \cap V^\star, 
	degree(a) = 2$
\end{enumerate}

@@

\subsection{Modeling \relonto{} in CP}
In order to model this problem we used the \minizinc{} language 
\cite{minizinc}, and the 
\chuffed{} solver \cite{chu2011improving} because it has a global constraint 
implemented for Steiner 
Tree Problems \cite{deuna2016steiner}.

Because attributes must be connected to exactly one node of the alignment 
graph, and that node will be in $\mathcal{D_{G_O}} \cap V^\star$ 
then the part of the problem between attribute nodes and the alignment graph is 
actually a matching problem. Each attribute must match exactly to one node of 
$\mathcal{D_{G_O}}$. Note that not all nodes in $\mathcal{D_{G_O}}$ must match 
to an attribute, as not all of them are part of $T^\star$.

@@

\begin{flalign}
	& \text{Minimize~~} w_{STP} + w_{ADIFF} - w_{PAT}
	%\sum_{e \in \mathcal{E}_\mathcal{O}}c_e * 
	%w_\mathcal{I}(e) + \sum_{e \in \mathcal{M}_\mathcal{O}^{s^\star}} c_e * 
    %		w_\mathcal{I}(e)
	\label{EQ:obj}&&\\
	&\text{such that~~}  \nonumber&& \\
	& \textit{steiner\_tree}(\{c_n| n \in \mathcal{V_O}\},\{c_e| e \in 
	\mathcal{E_O}\},\mathcal{G_O},w_\mathcal{O},w_\mathit{STP})  \label{EQ:stp} 
	&&\\
	&\forall d \in \mathcal{D_{G_O}}, degree(d) \leq 1 \label{EQ:deg1}&&\\
	&\forall d \in \mathcal{D_{G_O}}, c_d \Leftrightarrow degree(d) = 1 
	\label{EQ:deg2}&&\\
	&\forall a \in \mathcal{A}_{s^\star}, match[a] \in \{ d | (a,d)\in 
	\mathcal{M}_\mathcal{O}^{s^\star} \} \label{EQ:matchdom}&&\\
	& \textit{alldifferent}(match) \label{EQ:alld}&& \\
	& \forall a \in \mathcal{A}_{s^\star}, c_{match[a]} = \mathit{true} 
	\label{EQ:map}&&\\
	& w_{\mathit{ADIFF}} = \sum_{(a,d) \in \mathcal{M}_\mathcal{O}^{s^\star}} 
	w_\mathcal{I}(~(a,d)~) * \llbracket match[a] = d\rrbracket 
	\label{EQ:matchcost}  &&\\
	& \forall p \in \mathcal{P}, \big(\forall e\in edges(p), c_e = 
	\mathit{true} \big) \Leftrightarrow c_p \label{EQ:patt}&&\\
	& w_{PAT} = \sum_{p\in\mathcal{P}} \mathit{support}[p] * c_p 
	\label{EQ:pcost} 
\end{flalign}
